---
title: "Combing excel and Delimited data"
author: "Shohn Godboldt"
editor: visual
---
# Task 3: Combining Excel and Delimited Data

This section combines two datasets (red wine and white wine) and performs filtering and summarization.

## Read Excel Files

```{r}
# Load required libraries
library(tidyverse)
library(readxl)



##  White Wine Data

```{r}
# Read white wine data and add a "type" column
white_wine <- read_excel("white-wine.xlsx", sheet = 1) %>%
mutate(type = "white")
```

##  Red Wine Data

```{r}
# Read red wine data from CSV URL and add "type" column
red_wine <- read_delim(
"https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
delim = ";"
) %>%
mutate(type = "red")
```

## Combine, Filter, and Summarize

```{r}
# Combine red and white wine datasets into one
combined_wine <- bind_rows(white_wine, red_wine)

# Filter, sort, select, and summarize alcohol values by quality
combined_wine <- combined_wine %>%
filter(quality >= 5, alcohol < 13.2) %>%
arrange(desc(quality)) %>%
select(contains("acid"), alcohol, type, quality) %>%
group_by(quality) %>%
summarise(
alcohol_mean = mean(alcohol),
alcohol_sd = sd(alcohol)
)

# View result
combined_wine
```
---
